---
title: "Quantitative Methods, Assignment 10"
author: "rl2999@columbia.edu"
date: "December 12, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

CHIS_2017_teen <- 
  read_dta(file ="data/teen_2017_stata/TEEN.dta") 

#%>% as_factor()

```
<!-- You may use data of your own choosing for the assignment, but the dependent variable must be binary. -->

<!-- Estimate a multivariate logistic regression model that includes at least two independent variables. For this model describe: -->
<!-- The rationale for your hypothesis -->
<!-- Your interpretation of the regression model -->
<!-- Interpret the regression coefficients: -->
<!-- Use predicted probability for different values of independent variable -->
<!-- Using odds ratios -->
<!-- For each independent variable test whether it is statistically significant at your chosen level of significance. -->
<!-- Test whether the regression model as a whole is statistically significant at your chosen level of significance. -->

# Rationale

# Interpretation of regression model

# Interpretation of coefficients

## Predicted probability for different values of independent variable

## Interpretation by odds ratios

# Independent variable significance tests
<!-- For each independent variable test whether it is statistically significant at your chosen level of significance. -->

# Regression model significance tests
<!-- Test whether the regression model as a whole is statistically significant at your chosen level of significance. -->

Our model is 
Vapes

‘QT17_E4’ [TE66] -

‘QT17_D8’ [TD34] -
People in my neighborhood are willing to help each other.

‘QT17_A8’ [TA6] -
In the past 3 years, how many times did you change schools, not counting for graduation?

‘QT17_D12’ [TE64] -
Do you feel safe in your neighborhood…

‘QT17_D8’ [TD34] -
People in my neighborhood are willing to help each other.

‘QT17_L2’ [TH8] - How true do you feel the next statements are about your school and things you might do there:
At my school, there is a teacher or some other adult...
Who really cares about me. Would you say this is...

```{r}

summary(CHIS_2017_teen$te66)
summary(CHIS_2017_teen$ta6_p1v2)

CHIS_2017_teen_mod1 <- 
  CHIS_2017_teen %>% 
  glm(
  formula = .$te66 ~ ta6_p1v2 + te64 + th8,
  family = "binomial"
)
summary(CHIS_2017_teen_mod1)
plot(CHIS_2017_teen_mod1$residuals)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
